Limitations on the Two-Infall Scenario in the Era of Large Stellar Age Catalogs

### 2025-03-26 src/tex/ms.tex
Updates from Overleaf

\todo{Observational evidence.} \citet{nissen_high-precision_2020} measured abundances and estimated ages for nearby Solar-type stars and reported two sequences in the age--metallicity distribution which could be explained by the two-infall model \todo{(rewrite)}. \citet{nataf_accurate_2024} also observe multiple sequences in the age--metallicity relation in their sample of $\sim400,000$ local subgiants, although the break between them is not as clean as in \citet{nissen_high-precision_2020}. \citet{chen_recent_2025} however argue that this trend can be explained by a smooth star formation history with inside-out growth: the older sequence consists of stars which have migrated from the inner galaxy, where the star formation rate peaked early, while the younger sequence was formed locally.

@@

\todo{Migration.} The two-infall model attempts to reproduce the full distribution of stellar abundances in the Solar neighborhood through a single, continuous evolutionary track. \todo{Talk about migration studies and why we don't need a single evolutionary track to explain the bimodality.}

\todo{Summary of secular evolution models which claim alpha-bimodality.} \citep{kubryk_evolution_2015}. \citep{sharma_chemical_2021} . \citep{chen_chemical_2023}.\citep{prantzos_origin_2023}.

@@

\todo{Goal of paper.}

@@

The infall rate is described by two successive, exponentially declining bursts in time. The first infall component induces the formation of the thick disk, and the second component produces the thin disk. At a given galactocentric radius $R_{\rm gal}$, the un-normalized form of the infall rate is
\begin{equation}
    \label{eq:twoinfall-ifr}
    f_{\rm in}(t|R_{\rm gal}) = e^{-t/\tau_1} + f_{2/1}(R_{\rm gal}) e^{-(t-t_{\rm max})/\tau_2},
\end{equation}
where $\tau_1$ and $\tau_2$ are the first and second infall timescales, respectively, $t_{\rm max}$ is the onset of the second infall and thus the time of maximum gas infall, and $f_{2/1}$ is the ratio of the second infall amplitude to the first. We calculate $f_{2/1}$ for each zone such that the resulting stellar density profile follows a two-component disk, with the surface density ratio of the thick and thin disks given by
\begin{equation}
    f_\Sigma(R) \equiv \frac{\Sigma_1(R)}{\Sigma_2(R)} = f_{\Sigma,0} e^{R(1/R_2 - 1/R_1)}.
\end{equation}
We adopt a thick disk scale radius of $R_1=2.0\,{\rm kpc}$, a thin disk scale radius of $R_2=2.5\,{\rm kpc}$, and a central surface density ratio of $f_{\Sigma,0}=0.27$ \citep{bland-hawthorn_galaxy_2016}. \todo{Compare with other two-infall papers, which generally have a higher thick-to-thin ratio.}

@@

\begin{figure}
    \centering
    \includegraphics{figures/ofe_feh_density.pdf}
    \caption{The density of stars in the [O/Fe]--[Fe/H] plane from multi-zone models with (a) $y/Z_\odot=1$ and (b) $y/Z_\odot=2$, and (c) from the APOGEE DR17 catalog. The curves in panels (a) and (b) plot the ISM abundance of the Solar annulus over time, and the alternating black and white segments mark time intervals of {1 Gyr}. Stars are restricted to the region defined by $7\leq R_{\rm gal}< 9\,{\rm kpc}$ and $0.5\leq|z|<1\,{\rm kpc}$ to highlight the $\alpha$-bimodality.}
    \label{fig:ofe-feh-density}
\end{figure}

@@

>>>>

\todo{Observational evidence other than $\alpha$-bimodality.} \citet{nissen_high-precision_2020} measured abundances and estimated ages for nearby Solar-type stars and reported two sequences in the age--metallicity distribution which could be explained by the two-infall model \todo{(rewrite)}. \citet{nataf_accurate_2024} also observe multiple sequences in the age--metallicity relation in their sample of $\sim400,000$ local subgiants, although the break between them is not as clean as in \citet{nissen_high-precision_2020}. \citet{chen_recent_2025} however argue that this trend can be explained by a smooth star formation history with inside-out growth: the older sequence consists of stars which have migrated from the inner galaxy, where the star formation rate peaked early, while the younger sequence was formed locally.

@@

\todo{Migration.} The two-infall model attempts to reproduce the full distribution of stellar abundances in the Solar neighborhood through a single, continuous evolutionary track. \citet{sellwood_radial_2002} first showed that transient spiral perturbations can induce large changes in the guiding radius of a star without kinematic heating. It is now understood that the stars that make up the Solar neighborhood are drawn from a wide range of birth radii \citep[e.g.,][]{schonrich_chemical_2009,frankel_measuring_2018,lehmann_probing_2024}. Some studies have attempted to derive stellar birth radii \citep[e.g.,][]{ratcliffe_unveiling_2023,lu_there_2024}, though such an endeavor requires assumptions about the evolution of the Milky Way's radial metallicity gradient. While the strength and speed of radial migration in the disk are not precisely measured, it is clear that a single chemical evolution track need not explain the entirety of the local observed abundance distribution.

\todo{Summary of secular evolution models which claim alpha-bimodality.} \citet{schonrich_chemical_2009}. \citet{kubryk_evolution_2015}. \citet{sharma_chemical_2021}, \citet{chen_chemical_2023}. \citet{prantzos_origin_2023}.

@@

\todo{Goal and summary of paper.}

@@

The infall rate is described by two successive, exponentially declining bursts in time. The first infall component induces the formation of the thick disk, and the second component produces the thin disk. At a given galactocentric radius $R_{\rm gal}$, the un-normalized form of the infall rate is
\begin{equation}
    \label{eq:twoinfall-ifr}
    f_{\rm in}(t|R_{\rm gal}) = e^{-t/\tau_1} + f_{2/1}(R_{\rm gal}) e^{-(t-t_{\rm max})/\tau_2},
\end{equation}
where $\tau_1$ and $\tau_2$ are the first and second infall timescales, respectively, $t_{\rm max}$ is the onset of the second infall and thus the time of maximum gas infall, and $f_{2/1}$ is the ratio of the second infall amplitude to the first. We calculate $f_{2/1}$ for each zone such that the resulting stellar density profile follows a two-component disk, with the surface density ratio of the thick and thin disks given by
\begin{equation}
    f_\Sigma(R) \equiv \frac{\Sigma_1(R)}{\Sigma_2(R)} = f_{\Sigma,0} e^{R(1/R_2 - 1/R_1)}.
\end{equation}
We adopt a thick disk scale radius of $R_1=2.0\,{\rm kpc}$, a thin disk scale radius of $R_2=2.5\,{\rm kpc}$, and a central surface density ratio of $f_{\Sigma,0}=0.27$ \citep{bland-hawthorn_galaxy_2016}. In the Solar neighborhood, this produces a thick-to-thin disk mass ratio of $f_\Sigma(R_\odot)=0.12$. By comparison, \citet{mackereth_age-metallicity_2017} finds $f_\Sigma(R_\odot)=0.18\pm0.05$ from APOGEE. \todo{Compare with other two-infall papers, which generally have a higher thick-to-thin ratio.} \citet{spitoni_galactic_2020} explores $f_\Sigma(R_\odot)\approx0.23-0.31$, \citet{palla_chemical_2020} assumes $f_\Sigma(R_\odot)=0.23$, \citet{spitoni_apogee_2021} derives $f_\Sigma(R_\odot)=0.18$ (again from APOGEE), \citet{spitoni_remind_2024} assumes $f_\Sigma(R_\odot)=0.4$.

@@

\begin{figure}
    \centering
    \includegraphics{figures/ofe_feh_density.pdf}
    \caption{The density of stars in the [O/Fe]--[Fe/H] plane from multi-zone models with (a) $y/Z_\odot=1$ and (b) $y/Z_\odot=2$, and (c) from the APOGEE DR17 catalog. The curves in panels (a) and (b) plot the ISM abundance of the Solar annulus over time, and the alternating black and white segments mark time intervals of {1 Gyr}. Stars are restricted to the region defined by $7\leq R_{\rm gal}< 9\,{\rm kpc}$ and $0.5\leq|z|<1\,{\rm kpc}$ to highlight the $\alpha$-bimodality in the Solar neighborhood. {\bf Key takeaway:} the two-infall model generically predicts a stellar overdensity at intermediate [O/Fe] and low metallicity, which is not observed in APOGEE.}
    \label{fig:ofe-feh-density}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\linewidth]{figures/abundance_evolution_yields.pdf}
    \caption{Caption}
    \label{fig:abundance-evolution-yields}
\end{figure}
